
\bfsection{2 Theory of heights}


\bfsubsection{Theorem 2.3}
\barquo{
We set $n = [K:\Q]$. Let $\{ \gro_1, \cdots , \gro_n \}$ be the integral basis of $O_K$. Then $\{ x\gro_1, \cdots , x\gro_n \}$ is a basis of V.
}
\begin{proof}
  There is a $c_{ij} \in \Z$ such that $x \gro_i = \sum_{j} c_{ij} \gro_i$. Set $C= (c_{ij}) \in M_n(\Z)$. Then $\det C = N_{K/\Q}(x) \neq 0 $, so we get $C \in GL_n(\Q)$. And we obtain the assertion.
\end{proof}

\bfsubsection{Proposition 2.5}
\barquo{
\[
h_K(x) \leq \sum_{\grs \in K(\C)} \log \left( \max_{1 \leq i \leq n} \{ \abs{x_i}_{\grs} \} \right).
\]
}
\begin{rem}
  \textblue{Misprint.} Add $1/[K:\Q]$ into the right.
\end{rem}

\bfsubsection{Proposition 2.6}
\barquo{
for any $x \in \ol{\Q}^n$.
}
\begin{rem}
  \textblue{Misprint.} Exclude the case $x = 0$.
\end{rem}




\bfsubsection{Proposition 2.8}
\barquo{
We consider two morphisms $\phi_1 \colon X \to \P^{m_1}$ and $\phi_2 \colon X \to \P^{m_2}$ over $\ol{\Q}$. If $\phi_1^* \calo_{\P^{m_1}} \cong \phi_2^* \calo_{\P^{m_2}}$, then there is a constant $C$ such that, for any $x \in X(\ol{\Q})$,
\[
\abs{ h_{\phi_1}(x) -  h_{\phi_2}(x) } \leq C.
\]
}
\begin{proof} ${}$
  Remark that $\calo_{\P^{m_1}}(1)$ is a Serre's twisted sheaf. See Bosch\cite{Bosch} 9.2/Definition 3. or Hartshorne\cite{ha} section 2.5 Adjacent to Proposition 5.12.
We set $L = \phi_1^* \calo_{\P^{m_1}}$ and set $k = \ol{\Q}$ and $\scrf = \calo_{\P^{m_1}}(1)$. Since $X$ is a projective variety over $k$ and $L$ is an invertible sheaf on $X$, so $H^0(X,L) = \Gamma(X,L)$ is a $k$-vector space of finite dimension. (See Hartshorne\cite{ha} section 2.5 Theorem 5.19. and Hartshorne\cite{ha} section 2.4 Prop 4.10)

Let $\{ t_0, \cdots , t_m \}$ be a basis of $H^0(X,L)$ and let $X_0, \cdots , X_{m_1}$ be the homogenous coordinates of $\P^{m_1}$. Note that each $X_i$ is a global section of $\scrf$. And we set $s_i = \phi_1^* X_i \in H^0(X,L)$, where $\phi_1^* X_i$ is the image of $X_i \in H^0(\P^m, \scrf)$ by the canonical map $\scrf \to {\phi_1}_* \phi_1^* \scrf$.
It follows from Hartshorne\cite{ha} section 2.7 Theorem 7.1 that $s_0, \cdots , s_{m_1}$ generate $L$. Because for any $x \in X$ the each germ $(s_i)_x$ is a linear combination of $(t_j)_x$, so $t_0, \cdots ,t_m$ generate $L$.

There is a morphism $\phi \colon X \to \P^m $ such that $L \cong \phi^* \calo_{\P^m}(1)$ and $s_i = \phi^* X_i$ under this isomorphism. See Hartshorne\cite{ha} section 2.7 Theorem 7.1(b). There is another explanation on what $\phi$ is. For any $x \in X$, we can consider the germ $(t_i)_x \in L_x$. Denote $(t_i)_x$ by $t_i(x)$. Since $L$ is a line bundle, $L_x \cong \calo_{X,x}$.
Then we define the map $\phi \colon X \to \P^m$ by $\phi(x) = (t_0(x), \cdots , t_m(x))$. Note that there is a scalar ambiguity in choice of morphism $L_x \to k$. If $\forall i \; t_i(x) = 0$, then $(t_i)_x$ cannot generate $L_x$, which is a contradiction. Thus for any $x \in X$,
there is an index $i$ such that $t_i(x) \neq 0$.

Then, the rest of the proof is almost trivial.
\end{proof}


\bfsubsection{Theorem 2.9}
\barquo{
First, suppose that $L$ is globally generated.
}
\begin{rem}
  What "globally generated" means? We say $L$ is globally generated iff there is an exact sequence $\bigoplus_I O_X \to L \to 0$. Even if $L$ is an invertible sheaf, $L$ is not necessarily globally generated. For example, set $X = \P^m$, $L = \calo_{\P^m}(-1)$. Since $\Gamma(X,L) = 0$, $L$ is not globally generated.
\end{rem}


\bfsubsection{Theorem 2.9}
\barquo{
Let
\[
\phi_{\abs{L}} \colon X \to \P(H^0(X,L))
\]
be a morphism associated to the complete linear system $\abs{L}$. We set $h_L = h_{\phi_{\abs{L}}}$.
}
\begin{proof}
Note that we want to get $h_L \in \Func(X)/B(X)$, which is not contained in  $\Func(X)$.

What is a $\P(H^0(X,L))$? I think it is isomorphic to $\P^m$ by taking a basis of $H^0(X,L)$.

  Set $k = \ol{\Q}$. Since $L$ be a globally generated line bundle on $X$, there is a basis $s_0, \cdots , s_m$ of $H^0(X,L)$ which generate $L$. Then we get a map $\phi_L \colon X \to \P^m$ such that $\phi_L^* \calo_{\P^m}(1) \cong L$. We define $h_L$ by $h_L = h_{\phi_L}$.
\end{proof}


\bfsubsection{Theorem 2.9}
\barquo{
Then ${s_i \ts t_j} $ induces a morphism $\phi \colon X \to \P^N$ such that $\phi^*(O_{\P^N(1)}) \cong L_1 \ts L_2$.
}
\begin{rem}
  How $s_i \ts t_j \in H^0(X,L_1) \ts_k H^0(X,L_2)$ define an element of $H^0(X,L_1 \ts L_2)$? Let $\scrf$ be a presheaf defined by $\scrf(U) = \grG(U,L_1) \ts_{\calo_{X}(U)} \grG(U,L_2)$. Then there is a canonical morphism $\scrf \to L_1 \ts L_2$ since $L_1 \ts L_2$ is the sheafification of $\scrf$.
  So we can see $s_i \ts t_j \in H^0(X,L_1 \ts L_2)$.

  We denote the image of $s_i \ts t_l$ by $s_it_j \in H^0(X,L_1 \ts L_2)$. Why $\{ s_it_j \}$ generate $L_1 \ts L_2$? Take a stalk.
\end{rem}


\bfsubsection{Theorem 2.9}
\barquo{
tell us that $L \ts A^n$ is globally generated for any sufficiently large $n$.
}
\begin{rem}
  The ampleness of $A$ implies that
  \begin{align*}
    \exists n_1 \st n \geq n_1 &\To L \ts A^n \; \text{is globally generated} \\
      \exists n_2 \st n \geq n_2 &\To  A^n \; \text{is globally generated}
  \end{align*}
  Then we set $n = \max_i \{ n_i\}$.
\end{rem}


\bfsubsection{Theorem 2.9}
\barquo{
Then, modulo $B(X)$, we have
\[
h_{f^*(L)} = h_{f^*(C) \ts f^*(C)^{-1}}
\]
}
\begin{rem}
  See G\"ortz Wedhorn\cite{GW} Remark 7.10.
\end{rem}


\bfsubsection{Theorem 2.9}
\barquo{
Then by (1), $h_L$ must be equal to $h_{L_1} - h_{L_2}$ modulo $B(X)$.
}
\begin{rem}
  Let $\grs \colon \{ \text{line bundles }\} \to \Func(X)/B(X)$ be a map which satisfies the properties (1), (2), (3). By (3), for globally generated line bundle $L$, we get $\grs_L = h_L$. Because $\grG(X,O_X) = k$, we obtain $\grs_{O_X} = 0$. Thus (1) implies that $\grs_L = h_L$ for general line bundle $L$.
\end{rem}


\bfsubsection{Proposition 2.10}
\barquo{
Let $B$ be the Zariski closed subset of $X$ defined by the ideal sheaf
\[
\Im (H^0(X,L) \ts L^{-1} \to O_X).
\]
}
\begin{rem}
  What is the morphism $H^0(X,L) \ts L^{-1} \to O_X$? Note that there is a canonical morphism $f^* f_* L \to L$ where $f \colon X \to \Spec k$ is a $k$-scheme structure. Note that $f_* L$ is isomorphic to $\wt{H^0(X,L)}$. We denote this canonical morphism $f^* f_* L \to L$ by
  \[
  H^0(X,L) \ts O_X \to L.
  \]
  This is surjective if $L$ is globally generated.

  In general, we define $V \ts_k O_X$ for $k$-module $V$, by setting
  \[
  V \ts_k O_k = f^{-1} \wt{V} \ts_{f^{-1}O_{\Spec k}} O_X = f^* \wt{V}.
  \]
Then we get $H^0(X,L) \ts L^{-1} \to O_X$ by tensoring $L^{-1}$.
\end{rem}


\bfsubsection{Proposition 2.10}
\barquo{
Then $\{  s s_i \}$ are linearly independent elements of $H^0(X,L)$.
}
\begin{rem}
  What are $s s_i$ ? Note that there is a canonical morphism
  \[
  H^0(X,L) \ts H^0(X,L_2) \to H^0(X,L \ts L_2) \cong H^0(X,L_1)
  \]
  Thus I guess $s s_i$ is the image of $s \ts s_i$.

  Moreover, why $s s_i$ are linearly independent? It suffices to show that the morphism of $k$-module
  \[
  s \colon H^0(X,L_2) \to H^0(X,L_1)
  \]
  is injective.

  We prepare the following lemma.
  \lem{
  Let $X$ be an integral scheme and let $L$ be a line bundle on $X$. Assume that $s \in H^0(X,L)$ is not zero. Then for any $x \in X$, $s_x \neq 0$ in $L_x$.
  }
  \begin{proof}
    Assume that there is a $z \in X$ such that $s_z = 0$. We want to show $s = 0 \in H^0(X,L)$. Since  $L$ is invertible, there is an open affine covering $X = \bigcup_{i \in I} U_i$ such that
    \[
    U_i = \Spec A_i , \; L|_{U_i} \cong \wt{A_i}
    \]
    On the other hand, $s_z = 0$ implies that there is an open subset $U \subset X$ such that $s|_U = 0$ and $z \in U$. Since $X$ is integral, $U \cap \Spec A_i \neq \emptyset$. Thus there is a $g_i \in A_i \sm \{ 0 \}$ such that $\emptyset \neq D(g_i) \subset U \cap \Spec A_i$. Then $s|_{D(g_i)} = 0$ in $\grG(D(g_i), L) \cong {A_i}_{g_i}$.
    Note that each $A_i$ is an integral domain because $X$ is integral. Thus we get $\forall i \; s|_{U_i}=0$ because $A_i \to {A_i}_{g_i}$ is injective. It follows from the sheaf axiom that $s=0 \in \grG(X,L)$.
  \end{proof}

  Then, we can prove the injectivity of $s \colon O_X \to L$. First, by the lemma, $0 \to O_X \xrightarrow{s} L$ is exact. Since $L_2$ is flat, $0 \to L_2 \xrightarrow{s} L_1$ is exact. Since global section is left exact, we get $0 \to H^0(X,L_2) \xrightarrow{s} H^0(X,L_1)$ is exact.
\end{rem}



\bfsubsection{Proposition 2.10}
\barquo{
Let $s_1, \cdots , s_n$ be a basis of $H^0(X,L)$. $\cdots$

 Because $B = \setmid{x \in X}{s_1(x) = \cdots = s_n(x) = 0}$.
}
\begin{proof}
  Why $B = \setmid{x \in X}{s_1(x) = \cdots = s_n(x) = 0}$? I guess
  \begin{align*}
    B &= \Supp \Coker(H^0(X,L) \ts L^{-1} \to O_X) \\
    &= \Supp \Coker(H^0(X,L) \ts O_X \to L) \ts L^{-1}  &(\text{right exactness of tensor}) \\
    &= \Supp \Coker(H^0(X,L) \ts O_X \to L) \cap \Supp L^{-1} \\
    &= \Supp \Coker(H^0(X,L) \ts O_X \to L)
  \end{align*}
  Thus we get
  \begin{align*}
    x \in B &\iff \Coker(H^0(X,L) \ts O_X \to L)_x \neq 0 \\
    &\iff \forall s \in H^0(X,L) \; x_x \in \frakm_x L_x \\
    &\iff s(x) = 0 \\
    &\iff s_1(x) = \cdots = s_n(x) = 0
   \end{align*}
\end{proof}


\bfsubsection{Lemma 2.11}
\barquo{
Then the definition of $h^+$ gives $h^+(\tau (x)) = h^+(x)$.
}
\begin{rem}
  Note that $\ord_P(\tau(x)) = \ord_{\tau^{-1}(P)} (x)$ and the residue degree at $P_i$ is independent of $i$ since $K / \Q$ is a Galois extension. (See Yukie\cite{雪江N2} Number 2 Cor 1.3,27)
\end{rem}

\begin{comment}
\bfsubsection{Lemma 2.21}
\barquo{



LEMMA2.21. $\;$ Let $X$ be a projective variety, and let $x_0$ and $x_1$ be closed points of $X$. Then there is a projective curve $C$ on $X$ passing through $x_0$ and $x_1$.

PROOF. If $x_0 = x_1$, then the result is obvious, so we assume that $x_0 \neq x_1$. We show the lemma by induction on dimension of $X$. The dimension one case is obvious. Suppose $\dim X \geq 2$ and let $\pi \colon \tilde{X} \to X$ be the blowup of $X$ along the two points $x_0,x_1$. We embed $\tilde{X}$ into projective space $\P^N$. We take $N$ to be the minimum among such embeddings. For a general hyperplane $H \subset \P^N$,
$\tilde{Y} = \tilde{X} \cap H$ is irreducible and is not contatined in $\pi^{-1}(x_i)$.(See [Hartshorne section.2 theorem8.18])
Further, because $\dim H + \dim \pi^{-1}(x_i) \geq N$, $H \cap \pi^{-1}(x_i) \neq \emptyset$. (See [Hartshorne section.1 theorem 7.2]) If we set $Y = \pi(\tilde{Y}) \subset X$, then $Y$ is a $(\dim X - 1)$-dimensional subvariety of $X$ that contains $x_0$ and $x_1$. By the induction hypothesis, there is a projective curve $C$ on $Y$ that passes through $x_0$ and $x_1$.
This $C$ is a desired curve.
}
\begin{que} ${}$
  \begin{description}
    \item[(1)] 「If $x_0 = x_1$, then the result is obvious」なのはどうしてでしょうか。その点を通る多項式$f$をとって$X \cap V(f)$を考えるとふつうは一つ次元が下がりますが、これを繰り返せばよいということでしょうか。
    \item[(2)] 「We embed $\tilde{X}$ into projective space $\P^N$.」とありますが、$X$が射影的な多様体なら$\tilde{X}$も射影的というHartshorne section 2 命題7.16から従っているのでしょうか。
    \item[(3)] 「$\tilde{Y} = \tilde{X} \cap H$ is irreducible and is not contatined in $\pi^{-1}(x_i)$.」とあります。引用されているBertiniの定理（とHartshorne section 3 注意7.9.1）から既約性がわかるのはいいとして、「and is not contatined in $\pi^{-1}(x_i)$.」はなぜわかるのでしょうか。
    \item[(4)] 「Further, because $\dim H + \dim \pi^{-1}(x_i) \geq N$, $H \cap \pi^{-1}(x_i) \neq \emptyset$」とあります。前半から後半が射影次元定理で従うのはよいです。前半はなぜ判るのでしょうか。
    \item[(5)] 「If we set $Y = \pi(\tilde{Y}) \subset X$, then $Y$ is a $(\dim X - 1)$-dimensional subvariety of $X$ that contains $x_0$ and $x_1$.」とあります。なぜ次元がこうなるのでしょうか。
  \end{description}
\end{que}



\newpage


\bfsubsection{Theorem 2.23}
\barquo{
Theorem 2.23 (Theorem of the cube II)

Let $X,Y$, and $Z$ be geometrically irreducible algebraic varieties over $F$, let $x_0 \in X(F)$, $y_0 \in Y(F)$, and $z_0 \in Z(F)$ be closed points, and let $L$ be a line bundle on $X \times Y \times Z$. Assume that $X$ and $Y$ are projective over $F$. Then $L$ is trivial if and only if
\[
 L|_{\{x_0\} \times Y \times Z}, \; L|_{X \times \{y_0\} \times Z}, \; L|_{X \times Y \times \{z_0\}}
\]
are all trivial.


\textbf{proof of  theorem 2.23}
\begin{description}
  \item[Step 1] Let $p_{13}$ denote the projection $X \times Y \times Z \to X \times Z$ to the first and the last factor. If we show that $L|_{ \{x\} \times Y \times \{z\} }$ is trivial for every closed points $x \in X$ and $z \in Z$, then, by the seesaw theorem, we can find a line bundle $M$ on $X \times Z$ such that $p_{13}^* M \cong L$. Further,
  $L|_{X \times \{y_0\} \times Z}$ is trivial
  by the hypothesis, so $M$ is trivial. Thus $L$ is trivial and we will obtain the theorem.
  \item[Step 2] $\;$ For any closed point $x \in X$, lemma 2.21 tells us that there is a projective cureve $C$ on $X$ passing $x_0$ and $x_1$. Let $\tilde{C} \to C$ be the normalization of $C$. Suppose that we show Theorem 2.23 for $X= \tilde{C}$. Then $L|_{\tilde{C} \times Y \times Z}$ will be trivial,
  so $L|_{ \{x\} \times Y \times \{z\} }$ will be trivial for any closed point $z \in Z$.
  As $x$ is an arbitrary closed point of $X$, by Step 1, we will obtain Theorem 2.23 for general $X$.
  \item[Step 3] By Step 2, we may assume that $X$ is a smooth projective curve. To conclude the proof, we here use some properties of the Jacobian variety $J$ of $X$, which we explain in Section 2.8 below. By [Hartshorne"Algebraic Geometry" section3 -9.9], $L$ has degree zero on each fiber of $X \times (Y \times Z) \to Y \times Z$. Thus $L$ defines a morphism
  \[
  f \colon Y \times Z \to J.
  \]
  Because both $\{y_0\} \times Z$ and $Y \times \{z_0\}$ are mapped to $0$ by $f$, the rigidity lemma tells us that the whole $Y \times Z$ is also mapped to $0$ by $f$. Thus $L$ is trivial on $X \times Y \times Z$.
\end{description}
}

\begin{que} ${}$
  \begin{description}
    \item[(1)]「To conclude the proof, we here use some properties of the Jacobian variety $J$ of $X$,」とありますが、$X$のヤコビ多様体とはなんでしょうか。（定義がわかりません）そしてここで使う性質とは何でしょうか。
    \item[(2)] 「$L$ has degree zero on each fiber of $X \times (Y \times Z) \to Y \times Z$. 」とあります。これはどうしてでしょうか。
    Hartshorneの引用された定理を見ると、Hilbert多項式がからんでいるようなのですが、そこから先に進めません。
    \item[(3)] 「Thus $L$ defines a morphism $f \colon Y \times Z \to J.$」とありますが、これはなぜでしょうか。どういう射でしょうか。
  \end{description}
  以上です。
\end{que}
\end{comment}


\bfsubsection{Lemma 2.17}
\barquo{
Because $X$ is projective, $p$ is a closed morphism
}
\begin{rem}
  $X \to \Spec F$はprojectiveなのでproperであり、とくにuniversally closedである。したがって$p$は閉写像。
\end{rem}

\bfsubsection{Lemma 2.19}
\barquo{
As the natural homomorphism $p^* M \to L$ coincides on each fiber with the natural homomorphism $\cdots$ we obtain that $p^* M \cong L$.
}
\begin{rem}
$X,Y$を$F$スキームとする。このとき$L_1,L_2$が$X \tm_F Y$上のラインバンドルであり、任意の$y \in Y$に対して
\[
L_1 |_{X \tm \{y\}} \cong L_2 |_{X \tm \{y\}}
\]
であったとしても$L_1 \cong L_2$であるとは限らない。

例を挙げよう。$p \colon X \tm_F Y \to Y$を射影とし、$M$を$Y$上のラインバンドルとする。このとき$p^* M |_{X \tm \{y\}} \cong O_{X \tm Y}|_{X \tm \{y\} }$だが、必ずしも$p^* M \cong O_{X \tm Y}$ではない。たとえば$X = \Spec F$, $Y = \P_F^m$, $M = \calo_{\P}(1)$
とすればよい。
\end{rem}



\bfsubsection{Lemma 2.21}
\barquo{
Let $X$ be a projective variety, and let $x_0$ and $x_1$ be closed points of $X$. Then there is a projective curve $C$ on $X$ passing through $x_0$ and $x_1$.
}
\begin{proof}
  $x_0 = x_1$ならこの点を通る直線$f$に対して$X \cap V(f)$を考えることを繰り返し、次元を$1$ずつ下げることができるのであきらか。（既約性をどう保証するのか？）$x_0 \neq x_1$とする。$\dim X = 1$ならあきらかなので$\dim X \geq 2$とする。$\dim X$についての帰納法で示そう。

  $\pi \colon \wt{X} \to X$を$x_0, x_1$に沿う$X$のblowupとする。$X$が射影的なので$\wt{X}$も射影的。よって閉埋め込み$\wt{X} \to \P^N$がある。ここで$N$はそのような性質を満たす最小のものをとることにする。

  ここで$x_0, x_1$は閉点なので、各$i$について$x_i$を含み$x_{1-i}$を含まないようなアファイン近傍$U_i = \Spec A_i$がある。閉点$x_i$は$A_i$の極大イデアル$\frakm_i$に対応しているとしよう。このとき$\pi^{-1}(x_i)$の次元がどうなっているかを考えたい。$X$がsmoothなら$\pi^{-1}(x_i) \cong \P^{\dim X -1}$であるが、ここではそれは使えないので次のように計算する。
  \begin{align*}
    \pi^{-1}(x_i) &= \pi^{-1}(U_i) \tm_{U_i} x_i \\
    &\cong \Proj ( \oplus_{d \geq 0} \frakm_i^d ) \tm \Spec A_i / \frakm_i \\
    &\cong \Proj ( ( \oplus_{d \geq 0} \frakm_i^d ) \ts \Spec A_i / \frakm_i  )\\
    &\cong \Proj (  \oplus_{d \geq 0} \frakm_i^d / \frakm_i^{d+1} ) \\
    &\cong \Proj (\gr_{\frakm_i} A_i )
  \end{align*}
  である。$\dim X \geq 2$という仮定より
  \begin{align*}
    \dim \pi^{-1}(x_i) &= \dim \gr_{\frakm_i} A_i -1 \\
    &= \dim A_i - 1 \\
    &\geq 1
  \end{align*}
  がわかる。したがって$\P^N$のhyperplane $H$をとると、あきらかに$\dim H = N -1$なので$\dim H + \dim \pi^{-1}(x_i) \geq N$が従う。ゆえに射影次元定理(Hartshorne\cite{ha} section 1-7.2)よりすべての$i$について$H \cap \pi^{-1}(x_i) \neq \emptyset$が成り立つ。
  いまBertiniの定理により$H$が一般のhyperplaneとするとき$\wt{Y} = \wt{X} \cap H$は既約。よって$Y = \pi(\wt{Y})$とすれば$Y$は$X$の部分多様体であり、$H \cap \pi^{-1}(x_i) \neq \emptyset$より$Y$は$x_0,x_1$を含む。かつ$N$の最小性より(最小性を使わなくても一般の超平面で切っているのだから当然に思えるが)$\wt{Y} \neq \wt{X}$なので、
  $\dim Y \leq \dim \wt{Y} < \dim \wt{X} = \dim X$である。よってこの$Y$に帰納法が適用できて証明が回る。
\end{proof}



\bfsubsection{Cor 2.25}
\barquo{
Let $A$ be an abelian variety of dimension $g$ over $F$.
\begin{description}
  \item[(1)] For any integer $n$, $[n] \colon A \to A$ is a finite and flat morphism of degree $n^{2g}$.
  \item[(2)] The abelian group $A(\ol{F})$ is divisible, i.e., for any $x \in A(\ol{F})$ and for any positive integer $n$, there is a $y \in A(\ol{F})$ with $[n](y) = x$.
\end{description}
}
\begin{proof} ${}$
  \begin{description}
    \item[(1)] $L$をevenかつampleなline bundleとする。$[n]^* L = L^{n^2}$より$[n]^* L$もampleである。$[n]^*L$がampleということは、（閉とは限らない）埋め込み
    \[
    \psi_{\abs{\wt{L}}} \colon A \to \P(\grG (A, \wt{L})^{\vee}  )
    \]
    がある。ただし$\wt{L} = [n]^* L$であり、${}^{\vee}$は双対空間を表す。このとき次の図式は可換。
    \[
    \xymatrix{
    A  \ar[r]^-{\psi_{\abs{\wt{L}}}} & \P(\grG (A, \wt{L})^{\vee}  ) \\
    \Ker [n] \ar[u] \ar[r]^-{\psi_{\abs{ \Ker [n] }  }} & \P(\grG (\Ker [n], \wt{L}|_{\Ker [n] } )^{\vee}  ) \ar[u]
    }
    \]
    ここで$\wt{L}|_{\Ker [n] }$は自明なので$\Ker [n]$の既約性分への分解を$\Ker [n] = \coprod_{i \in I} P_i$とすると、各成分$P_i$の$\psi_{\abs{ \Ker [n] }} $による像は$\P^{0}(F)$に含まれる。つまり一点である。
    したがって、$\psi_{\abs{\wt{L}}}$は埋め込みなので各$P_i$は一点である。
    よって$\dim \Ker [n] =  0$である。

    また$[n]$はprojective varietyの間の射なのでprojectiveであり、とくに固有である。$\dim \Ker [n] =  0$であることをいま示したが、$[n]$は準同型なのですべての点のfiberの次元が等しい。ゆえに固有かつすべての点でのfiberの次元がゼロなので$[n]$はfiniteである。とくにfiberの次元が任意の点で等しいので$[n]$はflatである。

    次にdegreeについて考える。ここでのdegreeは交点数を用いて定義される。ample line bundle $L$について
    \[
    L^{\cdot g} = (L. \cdots . L)
    \]
    と定義する。右辺の$L$は$g$個ある。(Reference: 石井志保子「特異点入門」)交点数は多重線形なので
    \[
    ( [n]^* L)^{\cdot g} = (L^{n^2}. \cdots . L^{n^2}) = n^{2g} L^{\cdot g}
    \]
    が従う。よってdegreeの定義から$\deg [n] = n^{2g}$である。
    \item[(2)] $[n] \colon A \to A$は固有なのでその像$[n](A)$は閉部分多様体である。また$[n]$はflatなので
    \[
    \dim_0 A + \dim_0 \Ker [n] = \dim_0 [n](A)
    \]
    が成り立つ。よって$\dim \Ker [n] =  0$より$\dim A = \dim_0 A = \dim_0 [n](A) = \dim [n](A)$である。真閉部分多様体は次元が落ちるはずなので$A = [n](A)$がわかる。
  \end{description}
\end{proof}


\bfsubsection{Remark 2.27}
\barquo{
If $D$ is an effective Cartier divisor on an abelian variety $A$, then $\abs{2D}$ is base point free. In particular, $D$ is nef.
}
\begin{rem}
  まず用語について解説する。$D$がbase point freeとは、rational map
  \[
  \psi_{\abs{D}} \colon A \dashrightarrow \P ( \grG (A, D)^{\vee} )
  \]
  が$A$全体で定義されることである。言い換えれば、
  \[
  \setmid{x \in A}{ \forall  0 \neq s \in \grG(A,D) \; s(x)=0 } = \emptyset
  \]
  ということである。また$D$がnef(数値的正、ネフ)とは
\[
\forall C \text{ irreducible curve} \; (D.C) \geq 0
\]
(交点数がゼロ以上)として定義される。

さてeffectiveなCartier divisor $D$について、$2D$がbase point freeならば$D$がnefであることを確かめよう。$2D$がnefならあきらかに$D$もnefなので、はじめから$D$がbase point freeだと仮定して$D$がnefだといえばよい。

いま$D$はeffectiveなので既約曲線$C$が$C \not\subset D$である限り、$(D.C) \geq 0$となる。交点数は線形同値なもの同士を入れ替えても不変なので、どんな$C$についてもある$D$と線形同値な$D'$があって$C \not\subset D'$となることをいえばよい。

ハイリホーで示す。ある既約曲線$C$が存在して、すべての$D' \sim D$なる$D'$について$C \subset D'$であったとする。いま$\psi_{\abs{D}}$の定義から、任意の超曲面$H \subset \P ( \grG (A, D)^{\vee} )$に対して$D \sim \psi_{\abs{D}}^* H$である。$H$はある大域切断$s \in \grG(A,D)$により$H = \setmid{l}{l(s) = 0}$と表せる。
そこでこれを$H_s$とおく。このとき仮定から
\[
\forall s \; C \subset \psi_{\abs{D}}^* H_s
\]
であるが、$\psi_{\abs{D}}^* H_s = \setmid{x \in A}{s(x)=0}$であったため、$D$がbase point freeであったことより$C = \emptyset$となるしかない。これは矛盾である。
\end{rem}


\newpage

\bfsubsection{Remark 2.27}

\barquo{
For an $a \in A(\ol{F})$, we define the morphism $T_a \colon A_{\ol{F}} \to \Pic(A_{\ol{F}} )$ by
\[
T_a \colon x \mapsto x + a.
\]
For any given line bundle $L$ on $A_{\ol{F}}$, we define the map $\grl_L \colon A(\ol{F}) \to \Pic( A_{ \ol{F} } )$ by
\[
\grl \colon x \mapsto T_x^*L \ts L^{-1}.
\]

Cor (Theorem of Square)

As groups, the map $\grl_L$ is a homomorphism from $A(\ol{F})$ to $\Pic( A_{ \ol{F} } )$.
}

\barquo{
If $D$ is an effective Cartier divisor on an abelian variety $A$, then $\abs{2D}$ is base point free. In particular, $D$ is nef. Indeed, identifying $D$ with the corresponding Weil divisor, we write $D + a$ and $D-a$ for $T_a(D)$ and $T_{-a}(D)$ respectively. For any $x \in A(\ol{F})$, we choose a point $a \not\in \Supp (D-x) \cup \Supp(D +x)$. Then $x \not\in \Supp (D-a) \cup \Supp(D +a)$. Further, the theorem of the square implies that $(D-a) + (D+a) \sim 2D$.
Thus $\abs{2D}$ is base point free.
}




\begin{rem}
Hartshorne\cite{ha} section 2.6 命題6.11の、$X$が整分離的Noetherスキームで局所分解的なものとすると、Cartier divisorとWeil因子が同型になるということを踏まえて同一視をする。ここでAbelian多様体$A$はsmoothで、したがって局所環がregularであり、正則局所環はUFDであることから$A$は局所分解的であることに気を付ける。

閉点$x \in A(\ol{F})$が任意に与えられたとする。$x$を台に含まないような、$2D$と線形同値なeffective因子の存在をいえば、$\abs{2D}$がbase point freeであることが従う。そういう因子を構成しよう。

点$a \not\in \Supp(D-x) \cup \Supp(-D + x)$をとる。ここで因子$D-x$は像$T_{-x}(D)$を意味し、$-D + x$は逆元をとる写像を$i$として$T_x(i (D))$を意味する。このとき$a \not\in \Supp(D-x)$より$x \not\in \Supp(D-a)$であり、$a \not\in \Supp(-D+x)$より$x \not\in \Supp(D + a)$である。（本文には\textblue{誤植がある}）このときTheorem of squareにより
\begin{align*}
  (D-a) + (D+a) &= T_{-a}(D) + T_{a}(D) \\
  &= \grl_{D}(-a) + \grl_D(a) + 2D \\
  &= 2D
\end{align*}
だから$2D$と$(D+a) + (D-a)$は線形同値。$(D+a) + (D-a)$はあきらかにeffectiveなので$\abs{2D}$がbase point freeであることがいえた。



\end{rem}


\newpage

\bfsubsection{Cor 2.28}
\barquo{
Let $A$ be an abelian variety, and let $D$ be an effective Cartier divisor on $A$. We set $L = O_X(D)$. (In particular, $\abs{2D}$ is base point free) Then the following are equivalent.
\begin{description}
  \item[(1)] $\Ker (\grl_L)$ is a finite subgroup of $A(\ol{F})$.
  \item[(2)] A morphism $\Phi \colon A \to \P(H^0(A,L^2))$ associated to the complete linear system $\abs{2D}$ is a finite morphism onto its image.
  \item[(3)] $L$ is ample.
\end{description}

PROOF.
Properties (2) and (3) hold over $F$ if and only if those hold over $\ol{F}$, so we may assume that $F = \ol{F}$.

\begin{description}
  \item[(1) $\To$ (2)] Suppose that $\Phi$ maps a projective curve $C$ on $A$ to a single point. We are going to deduce a contradiction by showing that $\Ker (\grl_L)$ contains $C -C= \setmid{x_2 - x_1}{x_1, x_2 \in C(F)}$. We write $D = \sum_{i=1}^r a_i D_i$ with $a_i > 0$ and prime divisors $D_i$ on $A$.

  We claim that either $(C + x) \cap D_i = \emptyset $ or $C + x \subset D_i$ for any $i$ and for any $x \in A(F)$. Because $D_i$ is an effective Cartier divisor on an abelian variety, it is nef by Remark 2.27. In particular, $(D_i \cdot C) \geq 0$. It follows from
  \[
  0 = (D \cdot C) = a_1 (D_1 \cdot C) + \cdots + a_r (D_r \cdot C),
  \]
  that $(D_i \cdot C) = 0$. Further, because $C + x$ is algebraically equivalent to $C$, we have $(D_i \cdot C + x) =(D_i \cdot C)=0$. Thus we obtain the claim.

  Next, we claim that $D_i = D_i + x_1 - x_2$ for any $x_1, x_2 \in C(F)$. Indeed, if $y \in D_i$ is a closed point, then both $C - x_1 + y$ and $D_i$ contain $y$. Thus $C - x_1 + y \subset D_i$ by the above argument. It follows that $x_2 - x_1 + y \in D_i$, so $y \in D_i + x_1 - x_2$. Thus $D_i \subset D_i + x_1 - x_2$. By switching $x_1$ and $x_2$ in the above argument, we have $D_i \supset D_i + x_1 - x_2$. Hence we obtain the claim.

  We set $L_i = O_A(D_i)$. Then $L = L_1^{\ts a_1} \ts \cdots \ts  L_r^{\ts a_r}$. By the above claim, we have $T_{x_2 - x_1}^*(L_i) = L_i$ for each $i$. It follows that
  \begin{align*}
    T_{x_2 - x_1}^*(L) &= T_{x_2 - x_1}^*(L_1)^{\ts a_1} \ts \cdots \ts T_{x_2 - x_1}^*(L_r)^{\ts a_r} \\
    &= L_1^{\ts a_1} \ts \cdots \ts  L_r^{\ts a_r} \\
  &= L.
  \end{align*}
  Thus $x_2 - x_1 \in \Ker (\grl_L)$, and so $C -C \subset \Ker (\grl_L)$. This is a contradiction.

  \item[(2)$\To$(3)] Let $E$ be any coherent $O_A$-module on $A$. We are going to show that $E \ts L^{2n}$ is globally generated for any sufficiently large $2n$. We write $X$ for the image $\Phi(A)$ of $\Phi$. Because $\Phi_*(E)$ is coherent on $X$ and $O_X(1)$ is ample, there is an $n_0 > 0$ such that
  \[
  H^0(X, \Phi_*(E) \ts O_X(n) ) \ts O_X \to \Phi_*(E) \ts O_X(n)
  \]
  is surjective for any $n \geq n_0$ (See Hartshorne p.153). Pulling back by $\Phi$, we find that
  \[
  (*)  \quad  \quad    H^0(X, \Phi_*(E) \ts O_X(n) ) \ts O_A \to \Phi^* \Phi_*(E) \ts \Phi^* O_X(n)
  \]
  is also surjective.

  On the other hand, it follows from $\Phi^* O_X(n) \cong L^{2n}$ and the projection formula that $\Phi_*(E \ts L^{2n}) \cong \Phi_*(E) \ts O_X(n)$. Further, as $\Phi$ is a finite morphism, the canonical morphism $\Phi^* \Phi_*(E) \to E$ is surjective. Thus the surjectivity of $(*)$ implies the surjectivity of
  \[
  H^0(A, E \ts L^{2n} ) \ts O_A \to E \ts L^{2n}
  \]
  for every $n \geq n_0$. Thus $L^2$ is ample, so $L$ is ample.

  \item[(3)$\To$(1)] Let $p_i \colon A \tm A \to A$ denote the i-th projection. First, we note the equality
  \[
  \Ker \grl_L = \setmid{ x \in A(F)  }{ (m_A^* L^{-1} \ts p_1^* L)|_{A \tm \{ x \} } \text{ is trivial}  }.
  \]
  In particular, the seesaw theorem tells us that $\Ker \grl_L$ endowed with the reduced induced scheme structure is regarded as a closed subgroup scheme of $A$. We denote by $B$ the connected component of $\Ker(\grl_L)$ containing the identity. Then $B$ is an abelian subvariety of $A$. We are going to show that $\dim B = 0$.

  We set $L' := (m_A^* L^{-1} \ts p_1^* L \ts p_2^* L)|_{B \tm B}$ on $B \tm B$. Because $L'|_{B \tm \{x\}} = (m_A^* L^{-1} \ts p_1^* L)|_{B \tm \{x\}} \ts p_2^* L|_{B \tm \{x\}}$ is trivial for any $x \in B$
  and $L'|_{ \{0\} \tm B}$ is trivial, the seesaw theorem implies that $L'$ is trivial. Pulling back $L'$ by $([1]_B, [-1]_B) \colon B \to B \tm B$, we obtain that $L \ts [-1]_A^* L $ is trivial on $B$. On the other hand, because $L$ is ample on $A$, $L \ts [-1]_A^* L$ is ample on $A$. Thus $\dim B = 0$,
   and we conclude that $\Ker(\grl_L)$ is a finite set.
  \end{description}

}

\begin{comment}
\newpage

\begin{que} ${}$
\begin{description}
  \item[(1)] 「Properties (2) and (3) hold over $F$ if and only if those hold over $\ol{F}$」とあります。これは何故なのでしょうか。
  \item[(2)] (1)$\To$(2)の証明で、背理法（対偶）が使われているようです。(2)が成り立たず、$\Phi$がfiniteでなかったとして、$\Ker \grl_L$が無限集合であることを示す証明になっているように見えます。そこで$\Phi$がfiniteでなければある$x$があって、$\Phi^{-1}(x)$に含まれるprojectiveな曲線$C$があるということを使っているようですが、どうしてそういう$C$が存在するのでしょうか。

  $\Phi$のすべてのファイバーの次元が$0$だとすると、$\Phi$はproperなのでfiniteになります。よって$\Phi$がfiniteでないという仮定から、$\Phi$のある点でのファイバーの次元は$1$以上であり、したがってあるファイバーに含まれるような射影曲線$C$がとれる、という証明で問題ないですか？
  \item[(3)] 「Further, because $C + x$ is algebraically equivalent to $C$」とあります。algebraically equivalentとはどういう意味（定義）で、そしてどうしてそれが分かるのでしょうか。
  \item[(4)] (2)$\To$(3)の証明で、「$\Phi^* O_X(n) \cong L^{2n}$」とあります。これはどうしてでしょうか。
  \item[(5)] 「as $\Phi$ is a finite morphism, the canonical morphism $\Phi^* \Phi_*(E) \to E$ is surjective.」とあります。これはどうしてでしょうか。よく知られた事実なのでしょうか。
  \item[(6)] (3)$\To$(1)の証明で、「because $L$ is ample on $A$, $L \ts [-1]_A^* L$ is ample on $A$」とあります。これはどうしてでしょうか。$[-1]_A$が同型だからでしょうか？
  \item[(7)] 「Thus $\dim B = 0$」とあります。これは何故でしょうか。
  \item[(8)] 「 and we conclude that $\Ker(\grl_L)$ is a finite set.」とありますが、$\dim B = 0$からこれが従う理由がよくわかりません。どうしてでしょうか。
\end{description}

\end{que}
\end{comment}
